They're Clutching up! Team Momentum in Round-Based Esports

### 2022-07-31 presentation.qmd
typos, re-organize

---
title: "They're on Fire: Team Momentum in Round-Based First Person Shooter Esports"
author: 'Tony ElHabr'
format:
  revealjs:
    slide-number: true
    incremental: false
    theme: [dark, theme.scss]
execute: 
  include: false
  echo: false
  eval: false
---

@@

-   2021 Valorant Champions Tour (VCT) matches (qualifiers and majors)
-   5v5
-   First to 13 round wins
-   One team plays offense in 12 consecutive rounds before swapping sides
-   Must win by 2 cod_rounds
:::

@@

::: fragment
Often the defensive team will try to eliminate all o team members prior to making the defuse, but in some cases, they may try to "ninja" defuse.
:::

@@

```{r}
#| label: cod_heatmap_prep
cod_round_and_series_win_prop_by_side <- cod_rounds |> 
  group_by(pre_cumu_w, pre_cumu_l, is_offense) |> 
  summarize(
    n = n(),
    across(c(win_round, win_series), sum)
  ) |> 
  ungroup() |> 
  mutate(
    win_round_prop = win_round / n,
    win_series_prop = win_series / n
  )

cod_o_win_prop <- cod_round_and_series_win_prop_by_side |> 
  filter(is_offense) |> 
  mutate(
    diff_win_round_prop = win_round_prop - !!cod_o_win_prop_const
  )

cod_o_win_prop <- cod_o_win_prop |> 
  inner_join(
    cod_e_o_win_series_prop |> select(pre_cumu_w, pre_cumu_l, e_win_series_prop),
    by = c('pre_cumu_w', 'pre_cumu_l')
  ) |> 
  mutate(
    diff_win_series_prop = win_series_prop - e_win_series_prop
  )

@@

cod_heatmap_seq_labeller <- function(.df, .label, .num) {
  sprintf(
    '%s\n(%s/%s)',
    scales::percent(.df[[.label]], accuracy = 1),
    scales::comma(.df[[.num]]),
    scales::comma(.df$n)
  )
}

heatmap_text_layer <- function(..., .op, .df, .label, .num, .color, .threshold, .labeller) {
  .df_filt <- .df |> filter(`.op`(.df[[.label]], .threshold))
  if(nrow(.df_filt) == 0) {
    return(list(...))
  }
  list(
    ...,
    geom_text(
      data = .df_filt,
      color = .color,
      size = 11 / .pt,
      fontface = 'bold',
      aes(
        label = .labeller(.df = .df_filt, .label = .label, .num = .num),
        x = .data$pre_cumu_w + 0.5, 
        y = .data$pre_cumu_l + 0.5
      )
    )
  )
}

heatmap_text_layers <- function(
    ..., 
    .df,
    .threshold,
    .labeller,
    .f
) {
  list(
    ...,
    .f(.df = .df, .threshold = .threshold, .labeller = .labeller, .emphasize = 'low'),
    .f(.df = .df, .threshold = .threshold, .labeller = .labeller, .emphasize = 'high')
  )
}
```

```{r}
#| label: p_cod_o_round_win_prop
o_round_win_text_layer <- function(
    ..., 
    .df,
    .threshold,
    .labeller,
    .label = 'win_round_prop',
    .num = 'win_round',
    .emphasize
) {
  if(.emphasize == 'low') {
    op <- `<`
    .color <- blackish_background
  } else {
    op <- `>=`
    .color <- 'white'
  }
  list(
    ...,
    heatmap_text_layer(
      .df = .df,
      .op = op,
      .label = .label,
      .num = .num,
      .color = .color,
      .threshold = .threshold,
      .labeller = .labeller
    )
  )
}

p_cod_o_round_win_prop <- cod_o_win_prop |> 
  ggplot() +
  common_cod_heatmap_layers() +
  heatmap_tile(.fill = 'win_round_prop') +
  heatmap_text_layers(
    .df = cod_o_win_prop,
    .f = o_round_win_text_layer,
    .threshold = 0.52,
    .labeller = cod_heatmap_seq_labeller
  ) +
  ggsci::scale_fill_material('teal') +
  labs(
    title = 'Offensive Round Win %'
  )
p_cod_o_round_win_prop
save_plot_for_slide(p_cod_o_round_win_prop)
```

@@

```{r}
#| label: p_cod_o_round_win_prop_over_expected
cod_heatmap_div_labeller <- function(.df, .label, .num) {
  sprintf(
    '%+.1f\n(%s)', 
    100 * .df[[.label]], 
    scales::comma(.df$n)
  )
}

cod_max_diff_win_round_prop <- max(abs(cod_o_win_prop$diff_win_round_prop))
p_cod_o_round_win_prop_over_expected <- cod_o_win_prop |> 
  ggplot() +
  common_cod_heatmap_layers() +
  heatmap_tile(.fill = 'diff_win_round_prop') +
  heatmap_text_layers(
    .df = cod_o_win_prop, 
    .f = o_round_win_over_expected_text_layer,
    .threshold = -0.1,
    .labeller = cod_heatmap_div_labeller
  ) +
  scale_fill_gradient2(
    low = extract_ggsci_mid_color('purple'),
    high = extract_ggsci_mid_color('green'),
    mid = 'white',
    midpoint = 0,
    limits = c(
      -cod_max_diff_win_round_prop, 
      +cod_max_diff_win_round_prop
    )
  ) +
  labs(
    title = 'Offensive Round Win % Over Expected'
  )
p_cod_o_round_win_prop_over_expected
save_plot_for_slide(p_cod_o_round_win_prop_over_expected)
```

@@

```{r}
val_round_and_series_win_prop_by_side <- val_rounds |> 
  group_by(pre_cumu_w, pre_cumu_l, is_offense) |> 
  summarize(
    n = n(),
    across(c(win_round, win_series), sum)
  ) |> 
  ungroup() |> 
  mutate(
    win_round_prop = win_round / n,
    win_series_prop = win_series / n
  )

val_o_win_prop <- val_round_and_series_win_prop_by_side |> 
  filter(is_offense) |> 
  mutate(
    diff_win_round_prop = win_round_prop - !!val_o_win_prop_const
  )

drop_25th_round <- function(df) {
  df |> 
    filter(
      !(pre_cumu_w == 12 & pre_cumu_l == 12)
    )
}

@@

val_o_win_prop <- val_o_win_prop |> 
  inner_join(
    val_e_o_win_series_prop |> select(pre_cumu_w, pre_cumu_l, e_win_series_prop),
    by = c('pre_cumu_w', 'pre_cumu_l')
  ) |> 
  mutate(
    diff_win_series_prop = win_series_prop - e_win_series_prop
  ) |> 
  filter(pre_cumu_w <= 12, pre_cumu_l <= 12) |> 
  drop_25th_round()

# Joshua Broas, vlr.gg
common_val_labs <- function(...) {
  list(
    ...,
    labs(
      caption = '2021 VCT matches'
    )
  )
}

@@

val_heatmap_seq_labeller <- function(.df, .label, .num) {
  sprintf(
    '%s', 
    scales::percent(.df[[.label]], accuracy = 1)
  )
}

val_pre_swap_rounds <- val_o_win_prop |> 
  mutate(pre_cumu_wl = pre_cumu_w + pre_cumu_l) |> 
  filter(pre_cumu_wl == 12)

val_pre_swap_round_segments <- val_pre_swap_rounds |> 
  transmute(
    pre_cumu_w,
    across(pre_cumu_l, ~.x + 1),
    i = row_number(),
    n = 2
  ) |> 
  uncount(n) |> 
  group_by(i) |> 
  mutate(rn = row_number()) |> 
  ungroup() |> 
  mutate(
    across(pre_cumu_w, ~ifelse(rn == 2, .x + 1, .x))
  ) |> 
  add_row(
    pre_cumu_w = 13,
    pre_cumu_l = 0
  )

val_swap_path <- function(...) {
  list(
    ...,
    geom_path(
      data = val_pre_swap_round_segments,
      color = 'white',
      size = 1.5,
      aes(
        x = pre_cumu_w, 
        y = pre_cumu_l
      )
    )
  )
}

p_val_o_round_win_prop <- val_o_win_prop |> 
  ggplot() +
  common_val_heatmap_layers() +
  heatmap_tile(.fill = 'win_round_prop') +
  val_swap_path() +
  heatmap_text_layers(
    .df = val_o_win_prop, 
    .f = o_round_win_text_layer,
    .threshold = 0.8,
    .labeller = val_heatmap_seq_labeller
  ) +
  ggsci::scale_fill_material('teal') +
  labs(
    title = 'Offensive Round Win %'
  )
p_val_o_round_win_prop
save_plot_for_slide(p_val_o_round_win_prop)
```

@@

```{r}
#| label: p_cod_o_series_win_prop
o_series_win_text_layer <- function(
    ..., 
    .df,
    .threshold,
    .labeller, 
    .emphasize
) {
  if(.emphasize == 'low') {
    op <- `<`
    .color <- blackish_background
  } else {
    op <- `>=`
    .color <- 'white'
  }
  list(
    ...,
    heatmap_text_layer(
      .df = .df,
      .op = op,
      .label = 'win_series_prop',
      .num = 'win_series',
      .color = .color,
      .threshold = .threshold,
      .labeller = .labeller
    )
  )
}

p_cod_o_series_win_prop <- cod_o_win_prop |> 
  ggplot() +
  common_cod_heatmap_layers() +
  heatmap_tile(.fill = 'win_series_prop') +
  heatmap_text_layers(
    .df = cod_o_win_prop, 
    .f = o_series_win_text_layer,
    .threshold = 0.8,
    .labeller = cod_heatmap_seq_labeller
  ) +
  ggsci::scale_fill_material('orange') +
  labs(
    title = 'Offensive Series Win %'
  )
p_cod_o_series_win_prop
save_plot_for_slide(p_cod_o_series_win_prop)
```

@@

```{r}
#| label: p_cod_o_series_win_prop_over_expected
o_series_win_over_expected_text_layer <- function(
    ..., 
    .df,
    .threshold,
    .labeller, 
    .emphasize
) {
  if(.emphasize == 'low') {
    op <- `<`
    .color <- blackish_background
  } else {
    op <- `>=`
    .color <- 'white'
  }
  list(
    ...,
    heatmap_text_layer(
      .df = .df,
      .op = op,
      .label = 'diff_win_series_prop',
      .num = NULL,
      .color = .color,
      .threshold = .threshold,
      .labeller = .labeller
    )
  )
}

ggsci_cyan <- extract_ggsci_mid_color('cyan')
ggsci_brown <- extract_ggsci_mid_color('brown')
cod_max_diff_win_series_prop <- max(abs(cod_o_win_prop$diff_win_series_prop))
p_cod_o_series_win_prop_over_expected <- cod_o_win_prop |> 
  ggplot() +
  common_cod_heatmap_layers() +
  heatmap_tile(
    .fill = 'diff_win_series_prop'
  ) +
  heatmap_text_layers(
    .df = cod_o_win_prop, 
    .f = o_series_win_over_expected_text_layer,
    .threshold = 0.08,
    .labeller = cod_heatmap_div_labeller
  ) +
  scale_fill_gradient2(
    low = ggsci_brown,
    high = ggsci_cyan,
    mid = 'white',
    midpoint = 0,
    limits = c(
      -cod_max_diff_win_series_prop, 
      +cod_max_diff_win_series_prop
    )
  ) +
  labs(
    title = 'Offensive Series Win % Over Expected'
  )
p_cod_o_series_win_prop_over_expected
save_plot_for_slide(p_cod_o_series_win_prop_over_expected)
```

@@

```{r}
#| label: p_val_o_series_win_prop_over_expected
val_heatmap_div_labeller <- function(.df, .label, .num) {
  sprintf(
    '%+.1f', 
    100 * .df[[.label]]
  )
}

val_max_diff_win_series_prop <- max(abs(val_o_win_prop$diff_win_series_prop))
p_val_o_series_win_prop_over_expected <- val_o_win_prop |> 
  ggplot() +
  common_val_heatmap_layers() +
  heatmap_tile(
    .fill = 'diff_win_series_prop'
  ) +
  heatmap_text_layers(
    .df = val_o_win_prop, 
    .f = o_series_win_over_expected_text_layer,
    .threshold = 0.2,
    .labeller = val_heatmap_div_labeller
  ) +
  val_swap_path() +
  scale_fill_gradient2(
    low = ggsci_brown,
    high = ggsci_cyan,
    mid = 'white',
    limits = c(
      -val_max_diff_win_series_prop,
      +val_max_diff_win_series_prop
    ),
    midpoint = 0
  ) +
  labs(
    title = 'Offensive Series Win % Over Expected'
  )
p_val_o_series_win_prop_over_expected
save_plot_for_slide(p_val_o_series_win_prop_over_expected)
```

@@

p_filt_val_o_series_win_prop_over_expected <- filt_val_o_win_prop |> 
  ggplot() +
  common_val_heatmap_layers() +
  heatmap_tile(
    .fill = 'diff_win_series_prop'
  ) +
  # geom_abline(
  #   color = 'white',
  #   linetype = 2,
  #   aes(
  #     slope = 1,
  #     intercept = 0
  #   )
  # ) +
  heatmap_text_layers(
    .df = filt_val_o_win_prop, 
    .f = o_series_win_over_expected_text_layer,
    .threshold = 0.1,
    .labeller = val_heatmap_div_labeller
  ) +
  val_swap_path() +
  scale_fill_gradient2(
    low = ggsci_brown,
    high = ggsci_cyan,
    mid = 'white',
    limits = c(
      -filt_val_max_diff_win_series_prop,
      +filt_val_max_diff_win_series_prop
    ),
    midpoint = 0
  ) +
  labs(
    title = 'Offensive Series Win % Over Expected'
  )
p_filt_val_o_series_win_prop_over_expected
save_plot_for_slide(p_val_o_series_win_prop_over_expected)
```

@@

::: fragment
[Chance (2019)](https://content.iospress.com/articles/journal-of-sports-analytics/jsa200422) finds that probability of winning a game that the title series in the MLB and NHL is **not** constant (at 50%)
:::

@@

p_cod_b2b_round_win_prop <- cod_b2b_round_win_prop |> 
  ggplot() +
  b2b_round_win_prop_tile_layer(
    .df = cod_b2b_round_win_prop, 
    .round = 7, 
    .palette = 'pink', 
    .is_new = FALSE
  ) +
  b2b_round_win_prop_tile_layer(
    .df = cod_b2b_round_win_prop, 
    .round = 8, 
    .palette = 'light-blue'
  ) +
  b2b_round_win_prop_tile_layer(
    .df = cod_b2b_round_win_prop, 
    .round = 9, 
    .palette = 'light-green'
  ) +
  b2b_round_win_prop_tile_layer(
    .df = cod_b2b_round_win_prop, 
    .round = 10, 
    .palette = 'amber'
  ) +
  b2b_round_win_prop_tile_layer(
    .df = cod_b2b_round_win_prop, 
    .round = 11, 
    .palette = 'deep-orange'
  ) +
  b2b_round_win_prop_text_layers(.df = cod_b2b_round_win_prop) +
  scale_x_continuous(
    labels = 2:11,
    breaks = seq(2.5, 11.5, by = 1),
    expand = c(0, 0.5)
  ) +
  scale_y_continuous(
    labels = 7:11,
    breaks = seq(7.5, 11.5, by = 1),
    expand = c(0, 0)
  ) +
  theme(
    panel.grid.major = element_blank(),
    axis.title = element_text(hjust = 0.5),
    plot.title = ggtext::element_markdown(hjust = 0.5),
    axis.text = element_text(size = 16, face = 'bold')
  ) +
  common_cod_labs() +
  labs(
    title = 'Back-to-back round win %',
    x = 'Round',
    y = '# of cod_rounds in Complete Series'
  ) + 
  geom_text(
    data = tibble(
      label = 'Back-to-back wins occur most\nfrequently in the last round.',
      x = 9.5,
      y = 8.5
    ),
    size = 9 / .pt,
    fontface = 'bold',
    color = 'white',
    hjust = 0,
    vjust = 0.5,
    aes(x = x, y = y, label = label)
  ) +
  map(
    cod_b2b_round_win_prop_arws |> group_split(rn), 
    function(.x) {
      geom_curve(
        data = data.frame(),
        color = 'white',
        arrow = arw,
        aes(x = .x$x, y = .x$y, xend = .x$xend, yend = .x$yend),
        curvature = .x$curvature
      )
    }
  )
p_cod_b2b_round_win_prop
save_plot_for_slide(p_cod_b2b_round_win_prop)
# ## now accounting for number of cod_rounds that end up being played AND series winners...
# cod_rounds |> 
#   filter(win_series, n_rounds > 6L) |> 
#   summarize_b2b_round_wins(n_rounds)
# 
# ## and the losers...
# cod_rounds |> 
#   filter(!win_series) |> 
#   summarize_b2b_round_wins(n_rounds)
```

>>>>

---
title: "They're Clutching up! Team Momentum in Round-Based Esports"
author: 'Tony ElHabr'
format:
  revealjs:
    slide-number: true
    incremental: false
    theme: [dark, theme.scss]
execute: 
  include: false
  echo: false
  eval: false
---

@@

-   2021 Valorant Champions Tour (VCT) matches (qualifiers and majors)
-   5v5
-   First to 13 round wins
-   One team plays offense in 12 consecutive rounds before swapping sides
-   Must win by 2 rounds
:::

@@

::: fragment
Often the defensive team will try to eliminate all team members prior to making the defuse, but in some cases, they may try to "ninja" defuse.
:::

@@

```{r}
#| label: cod-funcs

@@

heatmap_text_layer <- function(..., .op, .df, .label, .num, .color, .threshold, .labeller, .size = 11 / .pt) {
  .df_filt <- .df |> filter(`.op`(.df[[.label]], .threshold))
  if(nrow(.df_filt) == 0) {
    return(list(...))
  }
  list(
    ...,
    geom_text(
      data = .df_filt,
      color = .color,
      size = .size,
      fontface = 'bold',
      aes(
        label = .labeller(.df = .df_filt, .label = .label, .num = .num),
        x = .data$pre_cumu_w + 0.5, 
        y = .data$pre_cumu_l + 0.5
      )
    )
  )
}

heatmap_text_layers <- function(
    ..., 
    .df,
    .threshold,
    .labeller,
    .f
) {
  list(
    ...,
    .f(.df = .df, .threshold = .threshold, .labeller = .labeller, .emphasize = 'low'),
    .f(.df = .df, .threshold = .threshold, .labeller = .labeller, .emphasize = 'high')
  )
}

cod_heatmap_seq_labeller <- function(.df, .label, .num) {
  sprintf(
    '%s\n(%s)',
    scales::percent(.df[[.label]], accuracy = 1), 
    scales::comma(.df$n)
  )
}

o_round_win_text_layer <- function(
    ..., 
    .df,
    .threshold,
    .labeller,
    .label = 'win_round_prop',
    .num = 'win_round',
    .emphasize
) {
  if(.emphasize == 'low') {
    op <- `<`
    .color <- blackish_background
  } else {
    op <- `>=`
    .color <- 'white'
  }
  list(
    ...,
    heatmap_text_layer(
      .df = .df,
      .op = op,
      .label = .label,
      .num = .num,
      .color = .color,
      .threshold = .threshold,
      .labeller = .labeller,
      .size = 11 / .pt
    )
  )
}

cod_heatmap_div_labeller <- function(.df, .label, .num) {
  sprintf(
    '%+.1f\n(%s)', 
    100 * .df[[.label]], 
    scales::comma(.df$n)
  )
}

ggsci_purple <- extract_ggsci_mid_color('purple')
ggsci_green <- extract_ggsci_mid_color('green')
ggsci_cyan <- extract_ggsci_mid_color('cyan')
ggsci_brown <- extract_ggsci_mid_color('brown')
```

```{r}
#| label: cod_rounds-calcs
cod_round_and_series_win_prop_by_side <- cod_rounds |> 
  group_by(pre_cumu_w, pre_cumu_l, is_offense) |> 
  summarize(
    n = n(),
    across(c(win_round, win_series), sum)
  ) |> 
  ungroup() |> 
  mutate(
    win_round_prop = win_round / n,
    win_series_prop = win_series / n
  )

cod_o_win_prop <- cod_round_and_series_win_prop_by_side |> 
  filter(is_offense) |> 
  mutate(
    diff_win_round_prop = win_round_prop - !!cod_o_win_prop_const
  ) |> 
  inner_join(
    cod_e_o_win_series_prop |> select(pre_cumu_w, pre_cumu_l, e_win_series_prop),
    by = c('pre_cumu_w', 'pre_cumu_l')
  ) |> 
  mutate(
    diff_win_series_prop = win_series_prop - e_win_series_prop
  )
```

```{r}
#| label: p_cod_o_round_win_prop
p_cod_o_round_win_prop <- cod_o_win_prop |> 
  ggplot() +
  common_cod_heatmap_layers() +
  heatmap_tile(.fill = 'win_round_prop') +
  heatmap_text_layers(
    .df = cod_o_win_prop,
    .f = o_round_win_text_layer,
    .threshold = 0.52,
    .labeller = cod_heatmap_seq_labeller
  ) +
  ggsci::scale_fill_material('teal') +
  labs(
    title = 'Offensive Round Win %'
  )
p_cod_o_round_win_prop
save_plot_for_slide(p_cod_o_round_win_prop)
```

@@

```{r}
#| label: p_cod_o_round_win_prop_over_expected
cod_max_diff_win_round_prop <- max(abs(cod_o_win_prop$diff_win_round_prop))
p_cod_o_round_win_prop_over_expected <- cod_o_win_prop |> 
  ggplot() +
  common_cod_heatmap_layers() +
  heatmap_tile(.fill = 'diff_win_round_prop') +
  heatmap_text_layers(
    .df = cod_o_win_prop, 
    .f = o_round_win_over_expected_text_layer,
    .threshold = -0.1,
    .labeller = cod_heatmap_div_labeller
  ) +
  scale_fill_gradient2(
    low = ggsci_purple,
    high = ggsci_green,
    mid = 'white',
    midpoint = 0,
    limits = c(
      -cod_max_diff_win_round_prop, 
      +cod_max_diff_win_round_prop
    )
  ) +
  labs(
    title = 'Offensive Round Win % Over Expected'
  )
p_cod_o_round_win_prop_over_expected
save_plot_for_slide(p_cod_o_round_win_prop_over_expected)
```

@@

```{r}
#| label: val-funcs
drop_25th_round <- function(df) {
  df |> 
    filter(
      !(pre_cumu_w == 12 & pre_cumu_l == 12)
    )
}

@@

## Joshua Broas, vlr.gg
common_val_labs <- function(...) {
  list(
    ...,
    labs(
      caption = '2021 VCT matches'
    )
  )
}

@@

val_heatmap_seq_labeller <- function(.df, .label, .num) {
  sprintf(
    '%s\n(%s)', 
    scales::percent(.df[[.label]], accuracy = 1), 
    scales::comma(.df$n)
  )
}

o_series_win_text_layer <- function(
    ..., 
    .df,
    .threshold,
    .labeller, 
    .emphasize
) {
  if(.emphasize == 'low') {
    op <- `<`
    .color <- blackish_background
  } else {
    op <- `>=`
    .color <- 'white'
  }
  list(
    ...,
    heatmap_text_layer(
      .df = .df,
      .op = op,
      .label = 'win_series_prop',
      .num = 'win_series',
      .color = .color,
      .threshold = .threshold,
      .labeller = .labeller,
      .size = 9 / .pt
    )
  )
}

val_heatmap_div_labeller <- function(.df, .label, .num) {
  sprintf(
    '%+.1f\n(%s)', 
    100 * .df[[.label]], 
    scales::comma(.df$n)
  )
}

o_series_win_over_expected_text_layer <- function(
    ..., 
    .df,
    .threshold,
    .labeller, 
    .emphasize
) {
  if(.emphasize == 'low') {
    op <- `<`
    .color <- blackish_background
  } else {
    op <- `>=`
    .color <- 'white'
  }
  list(
    ...,
    heatmap_text_layer(
      .df = .df,
      .op = op,
      .label = 'diff_win_series_prop',
      .num = NULL,
      .color = .color,
      .threshold = .threshold,
      .labeller = .labeller,
      .size = 9 / .pt
    )
  )
}
```

```{r}
#| label: val_rounds-calcs
val_round_and_series_win_prop_by_side <- val_rounds |> 
  group_by(pre_cumu_w, pre_cumu_l, is_offense) |> 
  summarize(
    n = n(),
    across(c(win_round, win_series), sum)
  ) |> 
  ungroup() |> 
  mutate(
    win_round_prop = win_round / n,
    win_series_prop = win_series / n
  )

val_o_win_prop <- val_round_and_series_win_prop_by_side |> 
  filter(is_offense) |> 
  mutate(
    diff_win_round_prop = win_round_prop - !!val_o_win_prop_const
  ) |> 
  inner_join(
    val_e_o_win_series_prop |> select(pre_cumu_w, pre_cumu_l, e_win_series_prop),
    by = c('pre_cumu_w', 'pre_cumu_l')
  ) |> 
  mutate(
    diff_win_series_prop = win_series_prop - e_win_series_prop
  ) |> 
  filter(pre_cumu_w <= 12, pre_cumu_l <= 12) |> 
  drop_25th_round()
```

```{r}
#| label: val_rounds_bothbuy-calcs
val_bothbuy_rounds <- val_rounds |> 
  filter(
    team_buy_type %in% c('semi-buy', 'full-buy'), 
    opponent_buy_type  %in% c('semi-buy', 'full-buy')
  )

val_bothbuy_o_win_prop_const <- val_bothbuy_rounds |> 
  filter(is_offense) |> 
  count(win_round) |> 
  mutate(prop = n / sum(n)) |> 
  filter(win_round) |> 
  pull(prop)

val_round_and_series_win_prop_by_side_bothbuy <- val_bothbuy_rounds |> 
  group_by(pre_cumu_w, pre_cumu_l, is_offense) |> 
  summarize(
    n = n(),
    across(c(win_round, win_series), sum)
  ) |> 
  ungroup() |> 
  mutate(
    win_round_prop = win_round / n,
    win_series_prop = win_series / n
  )

val_o_win_prop_bothbuy <- val_round_and_series_win_prop_by_side_bothbuy |> 
  filter(is_offense) |> 
  mutate(
    diff_win_round_prop = win_round_prop - !!val_bothbuy_o_win_prop_const
  ) |> 
  inner_join(
    val_e_o_win_series_prop |> select(pre_cumu_w, pre_cumu_l, e_win_series_prop),
    by = c('pre_cumu_w', 'pre_cumu_l')
  ) |> 
  mutate(
    diff_win_series_prop = win_series_prop - e_win_series_prop
  ) |> 
  filter(pre_cumu_w <= 12, pre_cumu_l <= 12) |> 
  drop_25th_round()
```

```{r}
#| label: val-funcs-2
val_pre_swap_rounds <- val_o_win_prop |> 
  mutate(pre_cumu_wl = pre_cumu_w + pre_cumu_l) |> 
  filter(pre_cumu_wl == 12)

val_pre_swap_round_segments <- val_pre_swap_rounds |> 
  transmute(
    pre_cumu_w,
    across(pre_cumu_l, ~.x),
    i = row_number(),
    n = 2
  ) |> 
  uncount(n) |> 
  group_by(i) |> 
  mutate(rn = row_number()) |> 
  ungroup() |> 
  mutate(
    across(pre_cumu_w, ~ifelse(rn == 2, .x + 1, .x))
  ) |> 
  add_row(
    pre_cumu_w = 12,
    pre_cumu_l = 0
  )

val_swap_path <- function(...) {
  list(
    ...,
    geom_path(
      data = val_pre_swap_round_segments,
      color = 'white',
      size = 1.5,
      aes(
        x = pre_cumu_w, 
        y = pre_cumu_l
      )
    )
  )
}
```

```{r}
#| label: p_val_o_round_win_prop
p_val_o_round_win_prop <- val_o_win_prop |> 
  ggplot() +
  common_val_heatmap_layers() +
  heatmap_tile(.fill = 'win_round_prop') +
  val_swap_path() +
  heatmap_text_layers(
    .df = val_o_win_prop, 
    .f = o_round_win_text_layer,
    .threshold = 0.8,
    .labeller = val_heatmap_seq_labeller
  ) +
  ggsci::scale_fill_material('teal') +
  labs(
    title = 'Offensive Round Win %'
  )
p_val_o_round_win_prop
save_plot_for_slide(p_val_o_round_win_prop)

p_val_o_round_win_prop_bothbuy <- val_o_win_prop_bothbuy |> 
  ggplot() +
  common_val_heatmap_layers() +
  heatmap_tile(.fill = 'win_round_prop') +
  val_swap_path() +
  heatmap_text_layers(
    .df = val_o_win_prop_bothbuy, 
    .f = o_round_win_text_layer,
    .threshold = 0.8,
    .labeller = val_heatmap_seq_labeller
  ) +
  ggsci::scale_fill_material('teal') +
  labs(
    title = 'Offensive Round Win %'
  )
p_val_o_round_win_prop_bothbuy
save_plot_for_slide(p_val_o_round_win_prop_bothbuy)
```

@@

```{r}
#| label: p_cod_o_series_win_prop
p_cod_o_series_win_prop <- cod_o_win_prop |> 
  ggplot() +
  common_cod_heatmap_layers() +
  heatmap_tile(.fill = 'win_series_prop') +
  heatmap_text_layers(
    .df = cod_o_win_prop, 
    .f = o_series_win_text_layer,
    .threshold = 0.8,
    .labeller = cod_heatmap_seq_labeller
  ) +
  ggsci::scale_fill_material('orange') +
  labs(
    title = 'Offensive Series Win %'
  )
p_cod_o_series_win_prop
save_plot_for_slide(p_cod_o_series_win_prop)
```

@@

```{r}
#| label: p_cod_o_series_win_prop_over_expected
cod_max_diff_win_series_prop <- max(abs(cod_o_win_prop$diff_win_series_prop))
p_cod_o_series_win_prop_over_expected <- cod_o_win_prop |> 
  ggplot() +
  common_cod_heatmap_layers() +
  heatmap_tile(
    .fill = 'diff_win_series_prop'
  ) +
  heatmap_text_layers(
    .df = cod_o_win_prop, 
    .f = o_series_win_over_expected_text_layer,
    .threshold = 0.08,
    .labeller = cod_heatmap_div_labeller
  ) +
  scale_fill_gradient2(
    low = ggsci_brown,
    high = ggsci_cyan,
    mid = 'white',
    midpoint = 0,
    limits = c(
      -cod_max_diff_win_series_prop, 
      +cod_max_diff_win_series_prop
    )
  ) +
  labs(
    title = 'Offensive Series Win % Over Expected'
  )
p_cod_o_series_win_prop_over_expected
save_plot_for_slide(p_cod_o_series_win_prop_over_expected)
```

@@

```{r}
#| label: p_val_o_series_win_prop_over_expected
val_max_diff_win_series_prop <- max(abs(val_o_win_prop$diff_win_series_prop))
p_val_o_series_win_prop_over_expected <- val_o_win_prop |> 
  ggplot() +
  common_val_heatmap_layers() +
  heatmap_tile(
    .fill = 'diff_win_series_prop'
  ) +
  heatmap_text_layers(
    .df = val_o_win_prop, 
    .f = o_series_win_over_expected_text_layer,
    .threshold = 0.2,
    .labeller = val_heatmap_div_labeller
  ) +
  val_swap_path() +
  scale_fill_gradient2(
    low = ggsci_brown,
    high = ggsci_cyan,
    mid = 'white',
    limits = c(
      -val_max_diff_win_series_prop,
      +val_max_diff_win_series_prop
    ),
    midpoint = 0
  ) +
  labs(
    title = 'Offensive Series Win % Over Expected'
  )
p_val_o_series_win_prop_over_expected
save_plot_for_slide(p_val_o_series_win_prop_over_expected)
```

@@

p_filt_val_o_series_win_prop_over_expected <- filt_val_o_win_prop |> 
  ggplot() +
  common_val_heatmap_layers() +
  heatmap_tile(
    .fill = 'diff_win_series_prop'
  ) +
  heatmap_text_layers(
    .df = filt_val_o_win_prop, 
    .f = o_series_win_over_expected_text_layer,
    .threshold = 0.1,
    .labeller = val_heatmap_div_labeller
  ) +
  val_swap_path() +
  scale_fill_gradient2(
    low = ggsci_brown,
    high = ggsci_cyan,
    mid = 'white',
    limits = c(
      -filt_val_max_diff_win_series_prop,
      +filt_val_max_diff_win_series_prop
    ),
    midpoint = 0
  ) +
  labs(
    title = 'Offensive Series Win % Over Expected'
  )
p_filt_val_o_series_win_prop_over_expected
save_plot_for_slide(p_filt_val_o_series_win_prop_over_expected)
```

![](plots/filt_val_o_series_win_prop_over_expected.png){fig-align="center"}

@@

::: fragment
[Chance (2019)](https://content.iospress.com/articles/journal-of-sports-analytics/jsa200422) finds that the probability of winning a game in the title series in the MLB and NHL is **not** constant (at 50%)
:::

@@

p_cod_b2b_round_win_prop <- cod_b2b_round_win_prop |> 
  ggplot() +
  b2b_round_win_prop_tile_layer(
    .df = cod_b2b_round_win_prop, 
    .round = 7, 
    .palette = 'pink', 
    .is_new = FALSE
  ) +
  b2b_round_win_prop_tile_layer(
    .df = cod_b2b_round_win_prop, 
    .round = 8, 
    .palette = 'light-blue'
  ) +
  b2b_round_win_prop_tile_layer(
    .df = cod_b2b_round_win_prop, 
    .round = 9, 
    .palette = 'light-green'
  ) +
  b2b_round_win_prop_tile_layer(
    .df = cod_b2b_round_win_prop, 
    .round = 10, 
    .palette = 'amber'
  ) +
  b2b_round_win_prop_tile_layer(
    .df = cod_b2b_round_win_prop, 
    .round = 11, 
    .palette = 'deep-orange'
  ) +
  b2b_round_win_prop_text_layers(.df = cod_b2b_round_win_prop) +
  scale_x_continuous(
    labels = 2:11,
    breaks = seq(2.5, 11.5, by = 1),
    expand = c(0, 0.5)
  ) +
  scale_y_continuous(
    labels = 7:11,
    breaks = seq(7.5, 11.5, by = 1),
    expand = c(0, 0)
  ) +
  theme(
    panel.grid.major = element_blank(),
    axis.title = element_text(hjust = 0.5),
    plot.title = ggtext::element_markdown(hjust = 0.5),
    axis.text = element_text(size = 16, face = 'bold')
  ) +
  common_cod_labs() +
  labs(
    title = 'Back-to-back round win %',
    x = 'Round',
    y = '# of rounds played in series'
  ) + 
  geom_text(
    data = tibble(
      label = 'Back-to-back wins occur most\nfrequently in the last round.',
      x = 9.5,
      y = 8.5
    ),
    size = 9 / .pt,
    fontface = 'bold',
    color = 'white',
    hjust = 0,
    vjust = 0.5,
    aes(x = x, y = y, label = label)
  ) +
  map(
    cod_b2b_round_win_prop_arws |> group_split(rn), 
    function(.x) {
      geom_curve(
        data = data.frame(),
        color = 'white',
        arrow = arw,
        aes(x = .x$x, y = .x$y, xend = .x$xend, yend = .x$yend),
        curvature = .x$curvature
      )
    }
  )
p_cod_b2b_round_win_prop
save_plot_for_slide(p_cod_b2b_round_win_prop)
# ## now accounting for number of cod_rounds that end up being played AND series winners...
# cod_rounds |> 
#   filter(win_series, n_rounds > 6L) |> 
#   summarize_b2b_round_wins(n_rounds)
# 
# ## and the losers...
# cod_rounds |> 
#   filter(!win_series) |> 
#   summarize_b2b_round_wins(n_rounds)
```
